Predicting the (non-cryptographic) PCG Pseudo-Random Number Generator In Practice

### 2020-06-01 article/main.tex
détails

\keywords{\todo{keyword}}

@@

Not all pseudo-random generators are of cryptographic strength. In some
applications, they are simply not necessary: to be used in Monte-Carlo
numerical simulations or generate random choices in games, a relaxed,
non-cryptographic notion of pseudo-randomness may be sufficient. This allows for
faster algorithms. For instance, \textsf{python} standard library's
\texttt{random} module uses the Mersenne Twister~\cite{MatsumotoN98}. The
\textsf{C} library that comes along \texttt{gcc} (the \texttt{glibc}) uses a
(poor) truncated linear congruential generator by default to implement the
\texttt{rand} function.

In the realm of non-cryptographic random generators, a PRG is deemed ``good
enough'' it is passes \emph{some} efficient statistical tests --- whereas the
cryptographic notion of pseudo-randomness asks that it passes \emph{any}
efficient test. There are \textit{de facto} statistical test suites; an initial
battery of randomness tests for RNGs was suggested by Knuth in the 1969 first
edition of The Art of Computer Programming. In 1996, Knuth's tests were then
supplanted by Marsaglia's \texttt{Diehard} tests. In 2007, L'Ecuyer proposed
the \texttt{TestU01}~\cite{LEcuyerS07} library, whose ``BigCrush'' test is
considered state-of-the-art by the relevant community, to the best of our
knowledge. In 2010, the NIST proposed its own statistical test suite (Special
Publication 800-22), to which improvements were later
suggested~\cite{ZhuMLZJ16}. We understand that the \texttt{PractRand} test suite
also has a good reputation.

@@

In most cases, it is fairly easy to see that a specific non-cryptographic PRG
does not meet the cryptographic notion of pseudo-randomness, and there are few
exceptions. Most are fairly easy to predict, meaning that after having observed
a prefix of the output, it is easy to produce the next ``pseudo-random''
bits. This makes a good source of exercises for cryptography students.

@@

\paragraph{Contribution.} We describe an algorithm that reconstructs the full
internal state of the strongest member of the \textsf{PCG} family. This allows
to predict the pseudo-random stream deterministically and clock the generator
backwards. The original seeds can also easily be reconstructed. The state
reconstruction algorithm is practical and we have executed it in practice. It
follows that predicting the output of the \textsf{PCG} should not be considered
challenging anymore.

@@

We implemented the algorithms in \textsf{C}, then asked the designer of the
\textsf{PCG} family to send us ``challenge'' pseudo-random streams ; we ran our
code and emailed back the (correct) seeds used to generate the challenge streams the next
day.

@@

Frieze \textit{et al}~\cite{Frieze} improved the efficiency of reconstruction
algorithms in simpler cases. They supposed that the multiplier $a$ and the
modulus $2^k$ were known and used lattice-based techniques to recover a
truncated linear congruential generator more efficiently, with more truncated
bits.  more efficiently.

@@

Lemma~\ref{thm:close} tells us that the approximation of a geometric sequence
obtained by dropping least-significant bits cannot be arbitrarily far from a
lattice point which reveals $U_0 \bmod 2^k$. Therefore, we may possibly
reconstruct truncated geometric series by finding the lattice vector closest to
the approximation we have. This means solving instances of the well-known
\textsf{Closest Vector Problem} (CVP), a fundamental hard problem on
lattices. It is NP-hard, and all known algorithms are exponential in the
dimension of the lattice, yet they can be fairly practical up to dimension
$\approx 70$.

@@

Starting from $n=\lceil 122/6\rceil\rceil$, we computed the length of the
shortest vector of the lattice spanned by $G_{n, 128}$ for each successive $n$
until the condition holds. The \textsf{Shortest Vector Problem} is another
well-known lattice NP-hard problem; we used the (almost) off-the-shelf
\texttt{G6K} library~\cite{AlbrechtDHKPS19}, which gave results very quickly by
sieving. \texttt{fplll}~\cite{fplll} was too slow above dimension 50, in the
default settings.

>>>>

\keywords{Pseudo-random number generator, guess-and-determine attack, truncated congruential generator, euclidean lattices, closest vector problem, practical attack}

@@

Not all pseudo-random generators are of cryptographic strength. In some
applications, it is simply not necessary: to be used in Monte-Carlo numerical
simulations or generate random choices in games, a relaxed, non-cryptographic
notion of pseudo-randomness may be sufficient. This allows for faster
algorithms. For instance, \textsf{python} standard library's \texttt{random}
module uses the Mersenne Twister~\cite{MatsumotoN98}. The \textsf{C} library
that comes along \texttt{gcc} (the \texttt{glibc}) uses a (poor) truncated
linear congruential generator by default to implement the \texttt{rand}
function.

In the realm of non-cryptographic random generators, a PRG is deemed ``good
enough'' it is passes \emph{some} efficient statistical tests --- whereas the
cryptographic notion of pseudo-randomness asks that it passes \emph{any}
efficient test. There are \textit{de facto} statistical test suites; an initial
battery of randomness tests for RNGs was suggested by Knuth in the 1969 first
edition of \emph{The Art of Computer Programming}. In 1996, Knuth's tests were then
supplanted by Marsaglia's \texttt{Diehard} tests. In 2007, L'Ecuyer proposed
the \texttt{TestU01}~\cite{LEcuyerS07} library, whose ``BigCrush'' test is
considered state-of-the-art by the relevant community, to the best of our
knowledge. In 2010, the NIST proposed its own statistical test suite (Special
Publication 800-22), to which improvements were later
suggested~\cite{ZhuMLZJ16}. We understand that the \texttt{PractRand} test suite
also has a good reputation.

@@

In most cases, it is fairly easy to see that a given conventional PRG does not
meet the cryptographic notion of pseudo-randomness, and there are few
exceptions. Most are fairly easy to predict, meaning that after having observed
a prefix of the output, it is easy to produce the next ``pseudo-random''
bits. This makes a good source of exercises for cryptology students.

@@

\paragraph{Contribution.} We describe an algorithm that reconstructs the full
internal state of the strongest member of the \textsf{PCG} family. This allows
to predict the pseudo-random stream deterministically and clock the generator
backwards. The original seeds can also easily be reconstructed. The state
reconstruction algorithm is practical and we have executed it in practice. It
follows that predicting the output of the \textsf{PCG} should be considered
practically feasible.

@@

We implemented our algorithms, then asked the designer of the \textsf{PCG}
family to send us ``challenge'' pseudo-random streams ; we ran our code and
emailed back the (correct) seeds used to generate the challenge streams the next
day.

@@

Frieze \textit{et al}~\cite{Frieze} improved the efficiency of reconstruction
algorithms in simpler cases. They supposed that the multiplier $a$ and the
modulus $2^k$ were known and used lattice-based techniques to recover a
truncated linear congruential generator with more truncated bits.

@@

Lemma~\ref{thm:close} tells us that the approximation of a geometric sequence
obtained by dropping least-significant bits cannot be arbitrarily far from a
lattice point which reveals $U_0 \bmod 2^k$. Therefore, we may possibly
reconstruct truncated geometric series by finding the lattice vector closest to
the approximation we have. This means solving instances of the well-known
\textsc{Closest Vector Problem} (CVP), a fundamental hard problem on
lattices. It is NP-hard, and all known algorithms are exponential in the
dimension of the lattice, yet they can be fairly practical up to dimension
$\approx 70$.

@@

Starting from $n=\lceil 122/6\rceil\rceil$, we computed the length of the
shortest vector of the lattice spanned by $G_{n, 128}$ for each successive $n$
until the condition holds. The \textsc{Shortest Vector Problem} (SVP) is another
well-known lattice NP-hard problem; we used the (almost) off-the-shelf
\texttt{G6K} library~\cite{AlbrechtDHKPS19}, which gave results very quickly by
sieving. \texttt{fplll}~\cite{fplll} was too slow above dimension 50, in the
default settings.
